\textbf{quicR}:\ An R Library for Streamlined Data Handling of Real-Time Quaking Induced Conversion Assays

### 2025-05-30 man/manuscript/softwareX_submission.tex
Updated the manuscript with the reproducibility study.

            \subsubsection{Visualization}
                The ensuing goal of this package is visualizing RT-QuIC data. While there are many ways to represent these data, \textbf{quicR} includes two functions which provide a rapid assessment of the RT-QuIC results.
                \begin{enumerate}
                    \item \texttt{plate\_view()}: accepts the output from the get\_real() function and get\_sample\_locations() function. Makes an 8$\times$12 or 16$\times$24 faceted plot depending on if the plate has 96 or 384 wells. Each facet shows the real-time data of each well.
                    \item \textbf{plot\_metrics()}: requires the output of calculate\_metrics() and generates a boxplot of each sample's MPR, MS, RAF, and TtT.
                \end{enumerate}

@@

            \begin{table}[H]
                \centering
                \begin{tblr}{
                    cells     = {font = \fontsize{11pt}{11pt}\selectfont},
                    colspec   = {rccc}, 
                    rows      = {rowsep = 1pt},
                    row{1}    = {font=\bfseries},
                    rowhead   = 1,
                    width     = \textwidth
                }
                    \hline
                    \SetCell[c=1]{c}Feature / Benchmark & \textbf{quicR} & \textbf{QuICSeedR} & \textbf{rtquicR} \\ 
                    \hline
                    Available on CRAN & Yes & Yes & No \\
                    Number of input files required & 1 & 2 & 2 \\ 
                    Number of threshold methods & 2 & 3 & 4 \\
                    Number of normalization methods & 1 & 1 & 2 \\
                    Automated metadata entry & Yes & No & No \\
                    Automated sample locations & Yes & Yes & No \\
                    Allows empty data & Yes & Yes & No \\
                    Compatible with 384-well plates & Yes & Yes & No \\
                    % Actively maintained & Yes & No \\
                    Built-in bulk analysis support & No & Yes & Yes \\
                    Diagnostic analysis functionality & No & Yes & No \\
                    Incorporates dilution factors & Yes & No & Yes \\
                    Converts time units & No & Yes & Yes \\
                    Calculates metrics independently & Yes & No & Yes \\
                    Calculates metrics in bulk & Yes & Yes & No \\
                    Calcualates MPR & Yes & Yes & Yes \\
                    Calculates MS & Yes & Yes & No \\
                    Calculates TtT & Yes & Yes & Yes \\
                    Calculates RAF & Yes & Yes & No \\
                    Calculates area under the curve & No & No & Yes \\
                    Calculates SD50 & No & No & Yes \\
                    \hline
                    Plate view plot execution time (s) & 0.082 & 0.062 & NA \\
                    Metric plot execution time (s) & 0.009 & 0.005 & 0.007 \\
                    Metric calculations (MPR, MS, etc.) (s) & 0.355 & 0.018 & NA\\
                    MPR calculation (s) & 0.002 & NA & 0.028 \\
                    TtT calculation (s) & 0.001 & NA & 0.131 \\
                    Metadata parsing time (s) & 0.127 & 0.037 & NA \\
                    Raw data parsing time (s) & 0.069 & 0.050 & 0.191 \\
                    Bulk analysis (s/file) & 6.550 & 0.849 & 1.986 \\
                    \hline
                \end{tblr}
                \caption{Comparison of core features and execution times between \textbf{quicR}, QuICSeedR, and rtquicR. While QuICSeedR demonstrated the fastest overall execution times, \textbf{quicR} offered enhanced metadata handling, metric flexibility, and support for unprocessed instrument output files.\ rtquicR introduces additional metrics such as SD50 and area under the curve but requires pre-structured input and lacks support for 384-well microplate formats and automated metadata parsing. All benchmarks were performed using the same dataset and machine (Windows 11, 16GB RAM) in 50 replicates.}\label{tbl:benchmark}
            \end{table}

        \subsection{Reproducibility}
            To promote transparency and reproducibility, we provide a standardized benchmarking and data extraction script alongside the \textbf{quicR} package. This script processes raw RT-QuIC output files in MARS export format, computes key kinetic metrics (MPR, MS, TtT, RAF), and writes all intermediate and final results to CSV files. Users are instructed to run the script locally, generating three files (\texttt{<user>\_meta.csv}, \texttt{<user>\_data.csv}, and \texttt{<user>\_calcs.csv}) which were then compared across systems or submissions to ensure consistent performance.

            To further guard against system-dependent variability, the script includes the following measures:

            \begin{enumerate}
                % \item Sorting of input and output data by sample ID to enforce deterministic row order;
                \item Cross-platform user ID resolution to enable standardized filenames
                \item Fixed random seed initialization to control for any stochastic components
                \item Output of full session metadata via \texttt{sessionInfo()} to record package versions and platform details
                \item UTF-8 encoding of all CSV files for consistent parsing across operating systems.
            \end{enumerate}

            This reproducibility framework ensures that benchmark results, as reported in Table~\ref{tbl:benchmark}, can be independently validated and extended. The script, along with example data, is included in the package repository and supplementary materials.

@@

        \subsection{Example Code}
            \# Step 1: Install and load the package.\\
            install.packages(``quicR'')\\
            library(quicR)\\
            \# Step 2: Identify the raw file.\\
            file <- ``example.xlsx''\\
            \# Step 3: Extract the raw data.\\
            raw <- get\_real(file)[[1]]\\
            \# Step 4: Normalize the data against the background.\\
            normal <- normalize\_RFU(raw, transposed = FALSE)\\
            \# Step 5: Extract the metadata.\\
            meta <- organize\_tables(file) |> convert\_tables()\\
            \# Step 6: Get sample locations.\\
            locations <- get\_sample\_locations(file)\\
            \# Step 7: Create the analyzed data frame.\\
            analyzed <- calculate\_metrics(normal, meta)\\
            \# Step 8: Plot the analyzed data frame.\\
            plot\_metrics(analyzed)\\
            \# Step 9: Plot the plate view.\\
            plate\_view(raw, locations)\\

@@

    \section*{Acknowledgements}
        Special thanks to Beni Altmann at The Comprehensive R Archive Network (CRAN) for help during the submission process to CRAN.\ We thank Tiffany Wolf and Marc Schwabenlander for their support through the Minnesota Center for Prion Research and Outreach. We thank Miranda Huang and Kristin Bondo for testing the code. We would like to acknowledge Suzanne Stone and Sarah Gresch for maintaining lab operations. This work was funded by the Legislative-Citizen Commission on Minnesota Resources.

>>>>

            \subsubsection{Visualization}
                The ensuing goal of this package is visualizing RT-QuIC data. While there are many ways to represent these data, \textbf{quicR} includes two functions which provide a rapid assessment of the RT-QuIC results.
                \begin{enumerate}
                    \item \textbf{plate\_view()}: accepts the output from the get\_real() function and get\_sample\_locations() function. Makes an 8$\times$12 or 16$\times$24 faceted plot depending on if the plate has 96 or 384 wells. Each facet shows the real-time data of each well.
                    \item \textbf{plot\_metrics()}: requires the output of calculate\_metrics() and generates a boxplot of each sample's MPR, MS, RAF, and TtT.
                \end{enumerate}

@@

            \begin{table}[H]
                \centering
                \begin{tblr}{
                    cells     = {font = \fontsize{11pt}{11pt}\selectfont},
                    colspec   = {rccc}, 
                    rows      = {rowsep = 1pt},
                    row{1}    = {font=\bfseries},
                    rowhead   = 1,
                    width     = \textwidth
                }
                    \hline
                    \SetCell[c=1]{c}Feature / Benchmark & \textbf{quicR} & \textbf{QuICSeedR} & \textbf{rtquicR} \\ 
                    \hline
                    Available on CRAN & Yes & Yes & No \\
                    Number of input files required & 1 & 2 & 2 \\ 
                    Number of threshold methods & 2 & 3 & 4 \\
                    Number of normalization methods & 1 & 1 & 2 \\
                    Automated metadata entry & Yes & No & No \\
                    Automated sample locations & Yes & Yes & No \\
                    Allows empty data & Yes & Yes & No \\
                    Compatible with 384-well plates & Yes & Yes & No \\
                    % Actively maintained & Yes & No \\
                    Built-in bulk analysis support & No & Yes & Yes \\
                    Diagnostic analysis functionality & No & Yes & No \\
                    Incorporates dilution factors & Yes & No & Yes \\
                    Converts time units & No & Yes & Yes \\
                    Calculates metrics independently & Yes & No & Yes \\
                    Calculates metrics in bulk & Yes & Yes & No \\
                    Calculates MPR & Yes & Yes & Yes \\
                    Calculates MS & Yes & Yes & No \\
                    Calculates TtT & Yes & Yes & Yes \\
                    Calculates RAF & Yes & Yes & No \\
                    Calculates area under the curve & No & No & Yes \\
                    Calculates SD50 & No & No & Yes \\
                    \hline
                    Plate view plot execution time (s) & 0.082 & 0.062 & NA \\
                    Metric plot execution time (s) & 0.009 & 0.005 & 0.007 \\
                    Metric calculations (MPR, MS, etc.) (s) & 0.355 & 0.018 & NA\\
                    MPR calculation (s) & 0.002 & NA & 0.028 \\
                    TtT calculation (s) & 0.001 & NA & 0.131 \\
                    Metadata parsing time (s) & 0.127 & 0.037 & NA \\
                    Raw data parsing time (s) & 0.069 & 0.050 & 0.191 \\
                    Bulk analysis (s/file) & 6.550 & 0.849 & 1.986 \\
                    \hline
                \end{tblr}
                \caption{Comparison of core features and execution times between \textbf{quicR}, QuICSeedR, and rtquicR. While QuICSeedR demonstrated the fastest overall execution times, \textbf{quicR} offered enhanced metadata handling, metric flexibility, and support for unprocessed instrument output files.\ rtquicR introduces additional metrics such as SD50 and area under the curve but requires pre-structured input and lacks support for 384-well microplate formats and automated metadata parsing. All benchmarks were performed using the same dataset and machine (Windows 11, 16GB RAM) in 50 replicates.}\label{tbl:benchmark}
            \end{table}

        \subsection{Reproducibility}
            To evaluate the reproducibility of \textbf{quicR} across systems and users, we developed a standardized script that processes MARS output files and writes three key result CSV files: metadata, normalized fluorescence data, and calculated metrics as well as a plain-text file containing R session information (See Table \ref{tbl:comparisons}). This script was distributed, and users were only expected to source the script in the project folder. The files were compared between each user for inconsistencies. Numerical data were given a floating point tolerance of 1E\textsuperscript{-13}. Additionally, session summaries were parsed to extract R version, environment, and machine information. We did not observe any divergent results between any two users. A copy of the benchmarking script is available in the GitHub repository.

            \begin{table}[H]
                \centering
                \begin{tblr}{
                    cells     = {font = \fontsize{10pt}{10pt}\selectfont},
                    colspec   = {c c c c c}, 
                    rows      = {rowsep = 1pt},
                    row{1}    = {font=\bfseries\fontsize{10pt}{10pt}\selectfont},
                    rowhead   = 1,
                    width     = \textwidth
                }
                    \hline
                    User & {R\\Version} & {quicR\\Version} & Platform & OS\\
                    \hline
                    1 & 4.4.3 & 2.1.0 & x86\_64-w64-mingw32/x64 & Windows 11 x64 (build 26100) \\
                    2 & 4.4.1 & 2.1.0 & aarch64-apple-darwin20  & macOS 15.5 \\
                    3 & 4.4.3 & 2.1.0 & aarch64-apple-darwin20  & macOS Sequoia 15.5 \\
                    4 & 4.5.0 & 2.1.0 & aarch64-apple-darwin20  & macOS Sonoma 14.7.5 \\
                    5 & 4.4.3 & 2.1.0 & x86\_64-w64-mingw32/x64 & Windows 11 x64 (build 26100) \\
                    \hline
                \end{tblr}
                \caption{User session information. The script was run on both Windows and Mac machines with a variety of R versions. The quicR versions were kept identical to ensure reproducibility.}\label{tbl:comparisons}
            \end{table}

@@

        \subsection{Example Code}
            \texttt{\# Step 1: Install and load the package.\\
            install.packages(``quicR'')\\
            library(quicR)\\
            \# Step 2: Identify the raw file.\\
            file <- ``example.xlsx''\\
            \# Step 3: Extract the raw data.\\
            raw <- get\_real(file)[[1]]\\
            \# Step 4: Normalize the data against the background.\\
            normal <- normalize\_RFU(raw, transposed = FALSE)\\
            \# Step 5: Extract the metadata.\\
            meta <- organize\_tables(file) |> convert\_tables()\\
            \# Step 6: Get sample locations.\\
            locations <- get\_sample\_locations(file)\\
            \# Step 7: Create the analyzed data frame.\\
            analyzed <- calculate\_metrics(normal, meta)\\
            \# Step 8: Plot the analyzed data frame.\\
            plot\_metrics(analyzed)\\
            \# Step 9: Plot the plate view.\\
            plate\_view(raw, locations)\\}

@@

    \section*{Acknowledgements}
        Special thanks to Beni Altmann at The Comprehensive R Archive Network (CRAN) for help during the submission process to CRAN.\ We thank Tiffany Wolf and Marc Schwabenlander for their support through the Minnesota Center for Prion Research and Outreach. We thank Miranda Huang and Kristin Bondo for testing the code. We would like to acknowledge Suzanne Stone and Sarah Gresch for maintaining lab operations. We thank Sarah Gresch, Kendra Phelps, Laramie Lindsey, Miranda Huang, and Kristin Bondo for their participation in the reproducibility study. This work was funded by the Legislative-Citizen Commission on Minnesota Resources.
